---
title: "Fungus heatmaps"
author: "Evan Gorstein"
date: "2023-04-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Reference: https://davetang.org/muse/2018/05/15/making-a-heatmap-in-r-with-the-pheatmap-package/
```{r}
library(pheatmap)
library(grid)
library(tidyverse)
library(here)
```


## Load data 

First we read in the data
```{r}
fung_abundance = readRDS(here("data", "clean", "fung_abundance.RDS"))
samp_metadata = readRDS(here("data", "clean", "samp_metadata.RDS"))
```

## Aggregate to higher taxonomic level
We aggregate the counts to the genus level. Note that the "Unknown" value aggregates the counts for all the OTUs whose genus are unknown
```{r}
genus.df = fung_abundance %>% 
  mutate(Genus = fct_na_value_to_level(Genus, "Unknown")) %>%
  group_by(Genus) %>%
  summarise(across(starts_with("Samp"), sum)) 

#Convert to matrix with genus label as rowname
genuses = genus.df$Genus
genus.mat = as.matrix(genus.df[,-1])
rownames(genus.mat) = genuses
```

Let's check what most abundant genuses are (including "Unknown")
```{r}
sort(rowSums(genus.mat), dec = T) %>% head(10)
```

## Filter data
Now we do some filtering.

First we filter out rare genuses. We apply three different thresholds so that we can see three different versions of the heatmap
```{r}
#Filter to only look at genuses that have at least 100 counts in total (summing over samples) 
genus_mask = rowSums(genus.mat) >= 100
genus.filt = genus.mat[genus_mask,]
#151 out of 409 genuses remain

#Filter to only look at genuses that have at least 1,000 counts in total (summing over samples) 
genus_mask_harsh = rowSums(genus.mat) >= 1000
genus.filt.harsh = genus.mat[genus_mask_harsh,]
#69 out of 409 families remain
```


Now we filter to get rid of samples that have very few of the remaining genuses
```{r}
##Filter to only look at samples in which at least 21 of the remaining 409 genuses show up
samp_mask = colSums(genus.mat > 0) >= 21
genus.mat = genus.mat[,samp_mask]
#Gets rid of 8 samples!

##Filter to only look at samples in which at least 21 of the remaining 151 genuses show up
samp_mask = colSums(genus.filt > 0) >= 21
genus.filt = genus.filt[,samp_mask]
#Gets rid of 10 samples!

##Filter to only look at samples in which at least 21 of the remaining 69 genuses show up
samp_mask = colSums(genus.filt.harsh > 0) >= 21
genus.filt.harsh = genus.filt.harsh[,samp_mask]
#Gets rid of 17 samples!
```


## Normalize data
Now we normalize with log ratio using count of "Unknown" genus as the reference (the denominator in the ratio). In order to do this and avoid negative infinite values, have to get rid of 0's by adding pseudo count of 0.5
```{r}
genus.smooth = genus.mat + 0.5
genus.filt.smooth = genus.filt + 0.5
genus.filt.harsh.smooth = genus.filt.harsh + 0.5

genus.lr = t( log(t(genus.smooth) / genus.smooth["Unknown",]) )
genus.lr.filt = t( log(t(genus.filt.smooth) / genus.filt.smooth["Unknown",]) )
genus.lr.harsh = t( log(t(genus.filt.harsh.smooth) / genus.filt.harsh.smooth["Unknown",]) )

#Now that we've normalized, safe to get rid of unknown genus
genus.lr = genus.lr[rownames(genus.lr)!="Unknown",]
genus.lr.filt = genus.lr.filt[rownames(genus.lr.filt)!="Unknown",]
genus.lr.harsh = genus.lr.harsh[rownames(genus.lr.harsh)!="Unknown",]
```


## Create dataframe for annotating heat map with fumigation-status and week sample metadata

This code can probably be re-written in more simple/readable way but works for now
```{r}
never_fumigated = samp_metadata %>% 
  filter(Time == "Day_0" | Treatment == "Non-fumigated chipping grass") %>% 
  pull(samp_number)

recently = samp_metadata %>% 
  filter(Time == "Day_10" & Treatment != "Non-fumigated chipping grass") %>% 
  pull(samp_number)

more_than_a_month = samp_metadata %>% 
  filter(Time != "Day_0" & Time != "Day_10" & Treatment != "Non-fumigated chipping grass") %>% 
  pull(samp_number)

# function that convert sample number to column names
make_sample = function(x){
  if(x>=0 & x<=9){
    return(paste0("Samp00", x))
  } else if(x>=10 & x<=99){
    return(paste0("Samp0", x))
  } else {
    return(paste0("Samp", x))
  }
}

#Create column names based on column numbers
samp_never_fumigated = c(sapply(never_fumigated, make_sample))
samp_recently = c(sapply(recently, make_sample))
samp_more = c(sapply(more_than_a_month, make_sample))

all_samps = colnames(genus.df)[-1]


samp_annotation = data.frame(`fumigation status` = factor(case_when(
    all_samps %in% samp_never_fumigated ~ "never",
    all_samps %in% samp_recently ~ "recent",
    all_samps %in% samp_more ~ "past"
  ))) %>%
  mutate(Time = samp_metadata$Time)
row.names(samp_annotation) = all_samps
```


## Create heatmaps

```{r}
# function to make a heatmap.
# argument: df(matrix): data, annotation(matrix): annotation for colname, colname, rowname(boolean): 
# show column\row name or not. ylabel, xlabel,title(String): specify labels and titles.
create_heatmap = function(df, annotation = NULL, 
                          colname = FALSE, 
                          rowname = FALSE, 
                          xlabel = NULL, 
                          ylabel = NULL, 
                          title = NULL, 
                          clustercol = TRUE){
  setHook("grid.newpage", function() pushViewport(viewport(x=1,y=1,width=0.9, height=0.9, name="vp", just=c("right","top"))), action="prepend")
  heatmap = pheatmap(df, annotation_col = annotation, 
                     show_colnames = colname, 
                     show_rownames = rowname, 
                     main = title, angle_col = 0, 
                     cluster_cols = clustercol, 
                     fontsize_row = 3)
  setHook("grid.newpage", NULL, "replace")
  grid.text(xlabel, y=-0.07, gp=gpar(fontsize=16))
  grid.text(ylabel, x=-0.07, rot=90, gp=gpar(fontsize=16))
  return (heatmap)
}

pdf(here("plots", "fungal_heatmap1.pdf"))
plt1 = create_heatmap(genus.lr,
                      annotation = samp_annotation,
                      rowname = FALSE, 
                      title = "Log ratio normalized abundances", 
                      xlabel = "232 Samples", 
                      ylabel = "408 fungal genuses"
)
dev.off()

pdf(here("plots", "fungal_heatmap2.pdf"))
plt1.filt = create_heatmap(genus.lr.filt,
                           annotation = samp_annotation,
                           title = "Log ratio normalized abundances", 
                           xlabel = "230 Samples", 
                           ylabel = "150 most abundant fungal genuses"
)
dev.off()

pdf(here("plots", "fungal_heatmap3.pdf"))
plt1.harsh = create_heatmap(genus.lr.harsh,
                            annotation = samp_annotation,
                            title = "Log ratio normalized abundances", 
                            xlabel = "223 samples", 
                            ylabel = "68 most abundant fungal genuses"
)
dev.off()

```
