---
title: "Bacterial Heatmaps"
author: "Evan Gorstein"
date: "2023-04-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Reference: https://davetang.org/muse/2018/05/15/making-a-heatmap-in-r-with-the-pheatmap-package/
```{r}
library(pheatmap)
library(grid)
library(tidyverse)
library(here)
```


## Load data 

First we read in the data
```{r}
fung_abundance = readRDS(here("data", "clean", "fung_abundance.RDS"))
samp_metadata = readRDS(here("data", "clean", "samp_metadata.RDS"))
```

## Aggregate to higher taxonomic level
We aggregate the counts to the phylum level. Note that the "Unknown_Family" value aggregates the counts for all the OTUs whose families are unknown
```{r}
phylum.df = fung_abundance %>% 
  mutate(Phylum = fct_na_value_to_level(Phylum, "Unknown_Phylum")) %>% #Missing values included in Unknown_Family level
  group_by(Phylum) %>%
  summarise(across(starts_with("Samp"), sum))  

#Convert to matrix with family label as rownames 
phylum = phylum.df$Phylum
phylum.mat = as.matrix(phylum.df[,-1])
rownames(phylum.mat) = phylum

```

Let's check what most abundant family are (including "Unknown_Phylum")
```{r}
sort(rowSums(phylum.mat), dec = T) %>% head(10)
```

```{r}
# save the most abundant phylum to a file
write.table(sort(rowSums(phylum.mat), dec = T), file = here("data", "clean", "most_abundant_phylum_fungi.txt"))
```

Total number of OTUs in the data
```{r}
nrow(phylum.mat)

# sum of all the counts
sum(rowSums(phylum.mat))
```

## Filter data

Now we filter out rare families. We apply three different thresholds so that we can see three different versions of the heatmap

```{r}

#Filter to only look at families that have at least 1,000 counts in total (summing over samples) 
phylum_mask = rowSums(phylum.mat) >= 1000
phylum.filt = phylum.mat[phylum_mask,]
#153 out of 296 families remain

#Filter to only look at families that have at least 10,000 counts in total (summing over samples) 
phylum_mask_harsh = rowSums(phylum.mat) >= 10000
phylum.filt.harsh = phylum.mat[phylum_mask_harsh,]
#74 out of 296 families remain

```

```{r}
# save 2 filtered phylum only to a file
write.table(rowSums(phylum.filt), file = here("data", "clean", "most_abundant_phylum_fung_filt.txt"))
write.table(rowSums(phylum.filt.harsh), file = here("data", "clean", "most_abundant_phylum_fung_filt_harsh.txt"))
```

If we wanted to filter `fam.filt` to only include samples in which at least 50 of the remaining 153 families show up (removing a single sample), we'd run the following coder
```{r}
#samp_mask = colSums(fam.filt > 0) >= 50
#fam.filt = fam.filt[,samp_mask]
#Gets rid of a single sample (sample 173)
```

If we wanted to normalize by scaling by library size, we'd run the following code
```{r}
##Normalize by dividing by library sizes
#fam.ls = t( t(fam.mat) / colSums(fam.mat) )
#fam.ls.filt = t( t(fam.filt) / colSums(fam.filt) )
#fam.ls.harsh = t( t(fam.filt.harsh) / colSums(fam.filt.harsh) )
#Get rid of missing family 
#fam.ls.nm = fam.ls[-nrow(fam.ls),]
#fam.ls.filt.nm = fam.ls.filt[-nrow(fam.ls.filt),]
#fam.ls.harsh.nm = fam.ls.harsh[-nrow(fam.ls.harsh),]
```


## Normalization

Now we normalize with log ratio using the counts of "Unknown_Family" as the reference (the denominator in the ratio). In order to do this and avoid negative infinite values, have to get rid of 0's by adding pseudo count of 0.5
```{r}
phylum.smooth = phylum.mat + 0.5
phylum.filt.smooth = phylum.filt + 0.5
phylum.filt.harsh.smooth = phylum.filt.harsh + 0.5

phylum.lr = t( log(t(phylum.smooth) / phylum.smooth["Unknown_Phylum",]) )
phylum.lr.filt = t( log(t(phylum.filt.smooth) / phylum.filt.smooth["Unknown_Phylum",]) )
phylum.lr.harsh = t( log(t(phylum.filt.harsh.smooth) / phylum.filt.harsh.smooth["Unknown_Phylum",]) )

# #Now that we've normalized, safe to get rid of "Unknown_Family"
phylum.lr = phylum.lr[rownames(phylum.lr)!="Unknown_Phylum",]
phylum.lr.filt = phylum.lr.filt[rownames(phylum.lr.filt)!="Unknown_Phylum",]
phylum.lr.harsh = phylum.lr.harsh[rownames(phylum.lr.harsh)!="Unknown_Phylum",]
```


## Create dataframe for annotating heat map with fumigation-status and week sample metadata

This code can probably be re-written in more simple/readable way but works for now
```{r}
never_fumigated = samp_metadata %>% 
  filter(Time == "Day_0" | Treatment == "Non-fumigated chipping grass") %>% 
  pull(samp_number)

recently = samp_metadata %>% 
  filter(Time == "Day_10" & Treatment != "Non-fumigated chipping grass") %>% 
  pull(samp_number)

more_than_a_month = samp_metadata %>% 
  filter(Time != "Day_0" & Time != "Day_10" & Treatment != "Non-fumigated chipping grass") %>% 
  pull(samp_number)

# function that convert sample number to column names
make_sample = function(x){
  if(x>=0 & x<=9){
    return(paste0("Samp00", x))
  } else if(x>=10 & x<=99){
    return(paste0("Samp0", x))
  } else {
    return(paste0("Samp", x))
  }
}

#Create column names based on column numbers
samp_never_fumigated = c(sapply(never_fumigated, make_sample))
samp_recently = c(sapply(recently, make_sample))
samp_more = c(sapply(more_than_a_month, make_sample))

all_samps = colnames(phylum.mat)


samp_annotation = data.frame(`fumigation status` = factor(case_when(
    all_samps %in% samp_never_fumigated ~ "never",
    all_samps %in% samp_recently ~ "recent",
    all_samps %in% samp_more ~ "past"
  ))) 
row.names(samp_annotation) = all_samps
```

## Create heatmaps

```{r}
# function to make a heatmap.
# argument: df(matrix): data, annotation(matrix): annotation for colname, colname, rowname(boolean): 
# show column\row name or not. ylabel, xlabel,title(String): specify labels and titles.
create_heatmap = function(df, annotation = NULL, 
                          colname = FALSE, 
                          rowname = FALSE, 
                          xlabel = NULL, 
                          ylabel = NULL, 
                          title = NULL, 
                          clustercol = TRUE){
  setHook("grid.newpage", function() pushViewport(viewport(x=1,y=1,width=0.9, height=0.9, name="vp", just=c("right","top"))), action="prepend")
  heatmap = pheatmap(df, annotation_col = annotation, 
                     show_colnames = colname, 
                     show_rownames = rowname, 
                     main = title, angle_col = 0, 
                     cluster_cols = clustercol, 
                     fontsize_row = 3)
  setHook("grid.newpage", NULL, "replace")
  grid.text(xlabel, y=-0.07, gp=gpar(fontsize=16))
  grid.text(ylabel, x=-0.07, rot=90, gp=gpar(fontsize=16))
  return (heatmap)
}

pdf(here("plots", "fung_heatmap1.pdf"))
plt1 = create_heatmap(phylum.lr,
                      annotation = samp_annotation,
                      rowname = TRUE, 
                      title = "Log ratio normalized abundances", 
                      xlabel = "Samples 1-240", 
                      ylabel = "All 12 fungal phyla"
)
dev.off()

pdf(here("plots", "fung_heatmap2.pdf"))
plt1.filt = create_heatmap(phylum.lr.filt,
                           annotation = samp_annotation,
                           title = "Log ratio normalized abundances", 
                           rowname = TRUE,
                           xlabel = "Samples 1-240", 
                           ylabel = "10 most abundant fungal phyla"
)
dev.off()

pdf(here("plots", "fung_heatmap3.pdf"))
plt1.harsh = create_heatmap(phylum.lr.harsh,
                            annotation = samp_annotation,
                            rowname = TRUE,
                            title = "Log ratio normalized abundances", 
                            xlabel = "Samples 1-240", 
                            ylabel = "4 most abundant fungal phyla"
)
dev.off()

```

# Heatmap for combined avergae of several replicates for the days.

We now visualize the heatmap for the average of the 3 groups of samples. We will use the same function as above but with a few modifications.

Below we first create the dataset.

```{r}
#Create average of 3 groups of samples

# transpose the matrix
phylum.lr.t = as.data.frame(t(phylum.lr))

# add sample fumigation status and time to the dataframe.
phylum.lr.t$fumigation_status = samp_annotation$fumigation.status
phylum.lr.t$time = samp_metadata$Time

# group by fumigation status and calculate the average
otu.phylum.lr = phylum.lr.t[, -c((ncol(phylum.lr.t)-1):ncol(phylum.lr.t))]
groupannotations.phylum.lr = phylum.lr.t[, (ncol(phylum.lr.t)-1):ncol(phylum.lr.t)]

otu_data.phylum.lr = cbind(otu.phylum.lr, groupannotations.phylum.lr)

# group by fumigation status and time and calculate the average
otu_data.phylum.lr = otu_data.phylum.lr %>% 
  group_by(fumigation_status, time) %>%
  summarise(across(everything(), mean), na.rm = TRUE) %>%
   pivot_longer(cols = -c(fumigation_status, time), names_to = "OTU", values_to = "value")

 heatmapdata.phylum.lr = otu_data.phylum.lr %>% 
 unite("group", fumigation_status, time, sep = "_") %>%
   pivot_wider(names_from = group, values_from = value)

# remove OTU with name na.rm
heatmapdata.phylum.lr = heatmapdata.phylum.lr %>% 
  filter(OTU != "na.rm")

heatmapdata.phylum.lr = as.tibble(heatmapdata.phylum.lr)

all_otus.phylum.lr = heatmapdata.phylum.lr$OTU

heatmapdata.phylum.lr = heatmapdata.phylum.lr %>% 
   select(-OTU)

row.names(heatmapdata.phylum.lr) = all_otus.phylum.lr
```

```{r}
phylum.lr.filt.t = as.data.frame(t(phylum.lr.filt))

phylum.lr.filt.t$fumigation_status = samp_annotation$fumigation.status
phylum.lr.filt.t$time = samp_metadata$Time

otu.phylum.lr.filt = phylum.lr.filt.t[, -c((ncol(phylum.lr.filt.t)-1):ncol(phylum.lr.filt.t))]
groupannotations.phylum.lr.filt = phylum.lr.filt.t[, (ncol(phylum.lr.filt.t)-1):ncol(phylum.lr.filt.t)]

otu_data.phylum.lr.filt = cbind(otu.phylum.lr.filt, groupannotations.phylum.lr.filt)

otu_data.phylum.lr.filt = otu_data.phylum.lr.filt %>% 
  group_by(fumigation_status, time) %>%
  summarise(across(everything(), mean), na.rm = TRUE) %>%
   pivot_longer(cols = -c(fumigation_status, time), names_to = "OTU", values_to = "value")

heatmapdata.phylum.lr.filt = otu_data.phylum.lr.filt %>%
    unite("group", fumigation_status, time, sep = "_") %>%
    pivot_wider(names_from = group, values_from = value)

heatmapdata.phylum.lr.filt = heatmapdata.phylum.lr.filt %>%
  filter(OTU != "na.rm")

heatmapdata.phylum.lr.filt = as.tibble(heatmapdata.phylum.lr.filt)

all_otus.phylum.lr.filt = heatmapdata.phylum.lr.filt$OTU

heatmapdata.phylum.lr.filt = heatmapdata.phylum.lr.filt %>% 
  select(-OTU)

row.names(heatmapdata.phylum.lr.filt) = all_otus.phylum.lr.filt
```


```{r}
phylum.lr.harsh.t = as.data.frame(t(phylum.lr.harsh))

phylum.lr.harsh.t$fumigation_status = samp_annotation$fumigation.status
phylum.lr.harsh.t$time = samp_metadata$Time

otu.phylum.lr.harsh = phylum.lr.harsh.t[, -c((ncol(phylum.lr.harsh.t)-1):ncol(phylum.lr.harsh.t))]
groupannotations.phylum.lr.harsh = phylum.lr.harsh.t[, (ncol(phylum.lr.harsh.t)-1):ncol(phylum.lr.harsh.t)]

otu_data.phylum.lr.harsh = cbind(otu.phylum.lr.harsh, groupannotations.phylum.lr.harsh)

otu_data.phylum.lr.harsh = otu_data.phylum.lr.harsh %>% 
  group_by(fumigation_status, time) %>%
  summarise(across(everything(), mean), na.rm = TRUE) %>%
   pivot_longer(cols = -c(fumigation_status, time), names_to = "OTU", values_to = "value")

heatmapdata.phylum.lr.harsh = otu_data.phylum.lr.harsh %>%
    unite("group", fumigation_status, time, sep = "_") %>%
    pivot_wider(names_from = group, values_from = value)

heatmapdata.phylum.lr.harsh = heatmapdata.phylum.lr.harsh %>%
  filter(OTU != "na.rm")

heatmapdata.phylum.lr.harsh = as.tibble(heatmapdata.phylum.lr.harsh)

all_otus.phylum.lr.harsh = heatmapdata.phylum.lr.harsh$OTU

heatmapdata.phylum.lr.harsh = heatmapdata.phylum.lr.harsh %>% 
  select(-OTU)

row.names(heatmapdata.phylum.lr.harsh) = all_otus.phylum.lr.harsh
```

Sample annotation data.

```{r}
# make sample annotations for each fumigation and day combination
all_sample_names = colnames(heatmapdata.phylum.lr)

# Creating annotations
annotation_df = data.frame(sample = all_sample_names)

# add fumigation status, based on the first part before the underscore
annotation_df$fumigation_status = sapply(strsplit(all_sample_names, "_"), function(x) x[1])

rownames(annotation_df) = all_sample_names

annotation_df = annotation_df %>%
  select(-sample)
```

Now we save the heatmap data to csv file with a new column for OTU's average count in each fumigation status.

```{r}
# no filter
heatmapdata.phylum.lr = as.data.frame(heatmapdata.phylum.lr)
heatmapdata.phylum.lr$never_average = rowMeans(heatmapdata.phylum.lr[, c("never_Day_0", "never_Day_10", "never_Day_255", "never_Day_282", "never_Day_43", "never_Day_71")])
heatmapdata.phylum.lr$recent_average = rowMeans(heatmapdata.phylum.lr[, "recent_Day_10", drop=FALSE])
heatmapdata.phylum.lr$past_average = rowMeans(heatmapdata.phylum.lr[, c("past_Day_255", "past_Day_282", "past_Day_43", "past_Day_71")])
heatmapdata.phylum.lr$never_total = rowSums(heatmapdata.phylum.lr[, c("never_Day_0", "never_Day_10", "never_Day_255", "never_Day_282", "never_Day_43", "never_Day_71")])
heatmapdata.phylum.lr$recent_total = rowSums(heatmapdata.phylum.lr[, "recent_Day_10", drop=FALSE])
heatmapdata.phylum.lr$past_total = rowSums(heatmapdata.phylum.lr[, c("past_Day_255", "past_Day_282", "past_Day_43", "past_Day_71")])

write.csv(heatmapdata.phylum.lr, here("data", "clean", "fung_phylum_heatmapdata.csv"), row.names = TRUE)

# filter 1
heatmapdata.phylum.lr.filt = as.data.frame(heatmapdata.phylum.lr.filt)
heatmapdata.phylum.lr.filt$never_average = rowMeans(heatmapdata.phylum.lr.filt[, c("never_Day_0", "never_Day_10", "never_Day_255", "never_Day_282", "never_Day_43", "never_Day_71")])
heatmapdata.phylum.lr.filt$recent_average = rowMeans(heatmapdata.phylum.lr.filt[, "recent_Day_10", drop=FALSE])
heatmapdata.phylum.lr.filt$past_average = rowMeans(heatmapdata.phylum.lr.filt[, c("past_Day_255", "past_Day_282", "past_Day_43", "past_Day_71")])
heatmapdata.phylum.lr.filt$never_total = rowSums(heatmapdata.phylum.lr.filt[, c("never_Day_0", "never_Day_10", "never_Day_255", "never_Day_282", "never_Day_43", "never_Day_71")])
heatmapdata.phylum.lr.filt$recent_total = rowSums(heatmapdata.phylum.lr.filt[, "recent_Day_10", drop=FALSE])
heatmapdata.phylum.lr.filt$past_total = rowSums(heatmapdata.phylum.lr.filt[, c("past_Day_255", "past_Day_282", "past_Day_43", "past_Day_71")])

write.csv(heatmapdata.phylum.lr.filt, here("data", "clean", "fung_phylum_heatmapdata_filt.csv"), row.names = TRUE)

# filter 2
heatmapdata.phylum.lr.harsh = as.data.frame(heatmapdata.phylum.lr.harsh)
heatmapdata.phylum.lr.harsh$never_average = rowMeans(heatmapdata.phylum.lr.harsh[, c("never_Day_0", "never_Day_10", "never_Day_255", "never_Day_282", "never_Day_43", "never_Day_71")])
heatmapdata.phylum.lr.harsh$recent_average = rowMeans(heatmapdata.phylum.lr.harsh[, "recent_Day_10", drop=FALSE])
heatmapdata.phylum.lr.harsh$past_average = rowMeans(heatmapdata.phylum.lr.harsh[, c("past_Day_255", "past_Day_282", "past_Day_43", "past_Day_71")])
heatmapdata.phylum.lr.harsh$never_total = rowSums(heatmapdata.phylum.lr.harsh[, c("never_Day_0", "never_Day_10", "never_Day_255", "never_Day_282", "never_Day_43", "never_Day_71")])
heatmapdata.phylum.lr.harsh$recent_total = rowSums(heatmapdata.phylum.lr.harsh[, "recent_Day_10", drop=FALSE])
heatmapdata.phylum.lr.harsh$past_total = rowSums(heatmapdata.phylum.lr.harsh[, c("past_Day_255", "past_Day_282", "past_Day_43", "past_Day_71")])

write.csv(heatmapdata.phylum.lr.harsh, here("data", "clean", "fung_phylum_heatmapdata_filt_harsh.csv"), row.names = TRUE)

```

Plotting the heatmap.

```{r}
pdf(here("plots_new", "fung_heatmap1_avg.pdf"))
plt2 = create_heatmap(heatmapdata.phylum.lr,
                      annotation = annotation_df,
                      rowname = TRUE, 
                      title = "Log ratio normalized abundances", 
                      xlabel = "Samples grouped by gime",  
                      ylabel = "All 12 fungal phyla"
)
dev.off()

pdf(here("plots_new", "fung_heatmap2_avg.pdf"))
plt2.filt = create_heatmap(heatmapdata.phylum.lr.filt,
                           annotation = annotation_df,
                           rowname = TRUE,
                           title = "Log ratio normalized abundances", 
                           xlabel = "Samples grouped by time", 
                           ylabel = "10 most abundant fungal phyla"
)
dev.off()

pdf(here("plots_new", "fung_heatmap3_avg.pdf"))
plt2.harsh = create_heatmap(heatmapdata.phylum.lr.harsh,
                            annotation = annotation_df,
                            rowname = TRUE,
                            title = "Log ratio normalized abundances", 
                            xlabel = "Samples grouped by time", 
                            ylabel = "4 most abundant fungal phyla"
)
dev.off()
```


We also do the same on Genus level again.

## Family Level Aggreagate

## Aggregate to higher taxonomic level
We aggregate the counts to the phylum level. Note that the "Unknown_Family" value aggregates the counts for all the OTUs whose families are unknown
```{r}
genus.df = fung_abundance %>% 
  mutate(Genus = fct_na_value_to_level(Genus, "Unknown_Genus")) %>% #Missing values included in Unknown_Family level
  group_by(Genus) %>%
  summarise(across(starts_with("Samp"), sum))  

#Convert to matrix with family label as rownames 
genus = genus.df$Genus
genus.mat = as.matrix(genus.df[,-1])
rownames(genus.mat) = genus

```

Let's check what most abundant family are (including "Unknown_Phylum")
```{r}
sort(rowSums(genus.mat), dec = T) %>% head(10)
```


## Filter data

Now we filter out rare families. We apply three different thresholds so that we can see three different versions of the heatmap

```{r}

#Filter to only look at families that have at least 1,000 counts in total (summing over samples) 
genus_mask = rowSums(genus.mat) >= 1000
genus.filt = genus.mat[genus_mask,]
#153 out of 296 families remain

#Filter to only look at families that have at least 10,000 counts in total (summing over samples) 
genus_mask_harsh = rowSums(genus.mat) >= 10000
genus.filt.harsh = genus.mat[genus_mask_harsh,]
#74 out of 296 families remain

```

If we wanted to filter `fam.filt` to only include samples in which at least 50 of the remaining 153 families show up (removing a single sample), we'd run the following coder
```{r}
#samp_mask = colSums(fam.filt > 0) >= 50
#fam.filt = fam.filt[,samp_mask]
#Gets rid of a single sample (sample 173)
```

If we wanted to normalize by scaling by library size, we'd run the following code
```{r}
##Normalize by dividing by library sizes
#fam.ls = t( t(fam.mat) / colSums(fam.mat) )
#fam.ls.filt = t( t(fam.filt) / colSums(fam.filt) )
#fam.ls.harsh = t( t(fam.filt.harsh) / colSums(fam.filt.harsh) )
#Get rid of missing family 
#fam.ls.nm = fam.ls[-nrow(fam.ls),]
#fam.ls.filt.nm = fam.ls.filt[-nrow(fam.ls.filt),]
#fam.ls.harsh.nm = fam.ls.harsh[-nrow(fam.ls.harsh),]
```


## Normalization

Now we normalize with log ratio using the counts of "Unknown_Family" as the reference (the denominator in the ratio). In order to do this and avoid negative infinite values, have to get rid of 0's by adding pseudo count of 0.5
```{r}
genus.smooth = genus.mat + 0.5
genus.filt.smooth = genus.filt + 0.5
genus.filt.harsh.smooth = genus.filt.harsh + 0.5

genus.lr = t( log(t(genus.smooth) / genus.smooth["Unknown_Genus",]) )
genus.lr.filt = t( log(t(genus.filt.smooth) / genus.filt.smooth["Unknown_Genus",]) )
genus.lr.harsh = t( log(t(genus.filt.harsh.smooth) / genus.filt.harsh.smooth["Unknown_Genus",]) )

#Now that we've normalized, safe to get rid of "Unknown_Family"
genus.lr = genus.lr[rownames(genus.lr)!="Unknown_Genus",]
genus.lr.filt = genus.lr.filt[rownames(genus.lr.filt)!="Unknown_Genus",]
genus.lr.harsh = genus.lr.harsh[rownames(genus.lr.harsh)!="Unknown_Genus",]
```



## Heatmap for combined average of 3 groups of samples

We now visualize the heatmap for the average of the 3 groups of samples. We will use the same function as above but with a few modifications.

Below we first create the dataset.

```{r}
#Create average of 3 groups of samples

# transpose the matrix
genus.lr.t = as.data.frame(t(genus.lr))

# add sample fumigation status to the dataframe
genus.lr.t$fumigation_status = samp_annotation$fumigation.status
genus.lr.t$time = samp_metadata$Time

# group by fumigation status and calculate the average
otu.genus.lr = genus.lr.t[, -c((ncol(genus.lr.t)-1):ncol(genus.lr.t))]
groupannotations.genus.lr = genus.lr.t[, (ncol(genus.lr.t)-1):ncol(genus.lr.t)]

otu_data.genus.lr = cbind(otu.genus.lr, groupannotations.genus.lr)

# group by fumigation status and time and calculate the average
otu_data.genus.lr = otu_data.genus.lr %>% 
  group_by(fumigation_status, time) %>%
  summarise(across(everything(), mean), na.rm = TRUE) %>%
   pivot_longer(cols = -c(fumigation_status, time), names_to = "OTU", values_to = "value")

 heatmapdata.genus.lr = otu_data.genus.lr %>% 
 unite("group", fumigation_status, time, sep = "_") %>%
   pivot_wider(names_from = group, values_from = value)

# remove OTU with name na.rm
heatmapdata.genus.lr = heatmapdata.genus.lr %>% 
  filter(OTU != "na.rm")

heatmapdata.genus.lr = as.tibble(heatmapdata.genus.lr)

all_otus.genus.lr = heatmapdata.genus.lr$OTU

heatmapdata.genus.lr = heatmapdata.genus.lr %>% 
   select(-OTU)

row.names(heatmapdata.genus.lr) = all_otus.genus.lr
```

```{r}
genus.lr.filt.t = as.data.frame(t(genus.lr.filt))

genus.lr.filt.t$fumigation_status = samp_annotation$fumigation.status
genus.lr.filt.t$time = samp_metadata$Time

otu.genus.lr.filt = genus.lr.filt.t[, -c((ncol(genus.lr.filt.t)-1):ncol(genus.lr.filt.t))]
groupannotations.genus.lr.filt = genus.lr.filt.t[, (ncol(genus.lr.filt.t)-1):ncol(genus.lr.filt.t)]

otu_data.genus.lr.filt = cbind(otu.genus.lr.filt, groupannotations.genus.lr.filt )

otu_data.genus.lr.filt = otu_data.genus.lr.filt %>% 
  group_by(fumigation_status, time) %>%
  summarise(across(everything(), mean), na.rm = TRUE) %>%
   pivot_longer(cols = -c(fumigation_status, time), names_to = "OTU", values_to = "value")

heatmapdata.genus.lr.filt = otu_data.genus.lr.filt %>%
    unite("group", fumigation_status, time, sep = "_") %>%
    pivot_wider(names_from = group, values_from = value)

heatmapdata.genus.lr.filt = heatmapdata.genus.lr.filt %>%
  filter(OTU != "na.rm")

heatmapdata.genus.lr.filt = as.tibble(heatmapdata.genus.lr.filt)

all_otus.genus.lr.filt = heatmapdata.genus.lr.filt$OTU

heatmapdata.genus.lr.filt = heatmapdata.genus.lr.filt %>% 
  select(-OTU)

row.names(heatmapdata.genus.lr.filt) = all_otus.genus.lr.filt
```


```{r}
genus.lr.harsh.t = as.data.frame(t(genus.lr.harsh))

genus.lr.harsh.t$fumigation_status = samp_annotation$fumigation.status
genus.lr.harsh.t$time = samp_metadata$Time

otu.genus.lr.harsh = genus.lr.harsh.t[, -c((ncol(genus.lr.harsh.t)-1):ncol(genus.lr.harsh.t))]
groupannotations.genus.lr.harsh = genus.lr.harsh.t[, (ncol(genus.lr.harsh.t)-1):ncol(genus.lr.harsh.t)]

otu_data.genus.lr.harsh = cbind(otu.genus.lr.harsh, groupannotations.genus.lr.harsh)

otu_data.genus.lr.harsh = otu_data.genus.lr.harsh %>% 
  group_by(fumigation_status, time) %>%
  summarise(across(everything(), mean), na.rm = TRUE) %>%
   pivot_longer(cols = -c(fumigation_status, time), names_to = "OTU", values_to = "value")

heatmapdata.genus.lr.harsh = otu_data.genus.lr.harsh %>%
    unite("group", fumigation_status, time, sep = "_") %>%
    pivot_wider(names_from = group, values_from = value)

heatmapdata.genus.lr.harsh = heatmapdata.genus.lr.harsh %>%
  filter(OTU != "na.rm")

heatmapdata.genus.lr.harsh = as.tibble(heatmapdata.genus.lr.harsh)

all_otus.genus.lr.harsh = heatmapdata.genus.lr.harsh$OTU

heatmapdata.genus.lr.harsh = heatmapdata.genus.lr.harsh %>% 
  select(-OTU)

row.names(heatmapdata.genus.lr.harsh) = all_otus.genus.lr.harsh
```

Plotting the heatmap.

```{r}
pdf(here("plots_new", "fung_heatmap1_avg_genus.pdf"))
plt2 = create_heatmap(heatmapdata.genus.lr,
                      annotation = annotation_df,
                      rowname = TRUE, 
                      title = "Log ratio normalized abundances", 
                      xlabel = "Samples Grouped by Time", 
                      ylabel = "All 408 fungal genuses"
)
dev.off()

pdf(here("plots_new", "fung_heatmap2_avg_genus.pdf"))
plt2.filt = create_heatmap(heatmapdata.genus.lr.filt,
                           annotation = annotation_df,
                           rowname = TRUE,
                           title = "Log ratio normalized abundances", 
                           xlabel = "Samples Grouped by Time", 
                           ylabel = "68 most abundant fungal genuses"
)
dev.off()

pdf(here("plots_new", "fung_heatmap3_avg_genus.pdf"))
plt2.harsh = create_heatmap(heatmapdata.genus.lr.harsh,
                            annotation = annotation_df,
                            rowname = TRUE,
                            title = "Log ratio normalized abundances", 
                            xlabel = "Samples Grouped by Time", 
                            ylabel = "22 most abundant fungal genuses"
)
dev.off()
```