---
title: "Permanova Analysis on Microbiome Dataset"
author: "Tianyi Xu"
date: "2023-04-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(vegan)
library(ggplot2)
library(tidyverse)
library(here)
library(forcats)
library(compositions)
library(dplyr)
library(stringr)
```


## Background

We would like to perform PERMANOVA analysis on the microbiome dataset to see if fumigation status and treatment has any effect on the relative abundance of microbes in the soil.

The following analysis analyze the family level data, and also from the species level.

Main package used: https://search.r-project.org/CRAN/refmans/vegan/html/adonis.html

## PERMANOVA Analysis on Family OTUs

## Data Preparation

Read in the data and prepare the data for analysis

Read in the cleaned data
```{r}
samp_metadata = readRDS(here("data", "clean", "samp_metadata.RDS"))
bac_abundance = readRDS(here("data", "clean", "bac_abundance.RDS"))
```

The following chunk of code is to make a family abundance dataframe - we only look at the family level first.
Note that the "Unknown_Family" value aggregates the counts for all the OTUs whose families are unknown
```{r}
# transpose the data
fam.df = bac_abundance %>% 
  mutate(Family = fct_na_value_to_level(Family, "Unknown_Family")) %>% #Missing values included in Unknown_Family level
  group_by(Family) %>%
  summarise(across(starts_with("Samp"), sum))  

#Convert to matrix with family label as rownames 
fams = fam.df$Family
fam.mat = as.matrix(fam.df[,-1])
rownames(fam.mat) = fams

fam.mat = t(fam.mat)
```


The following chunk of code is to make a sample metadata dataframe so we have independent variables (sample metadatas such as fumigation status and treatment) in the permanova analysis.
```{r}
sample_metadata = tibble(samp_metadata)
samps = rownames(fam.mat)
fam.mat = t(fam.mat)
rownames(sample_metadata) = samps

sample_metadata = samp_metadata %>%
  mutate(fumigation = case_when(
    Time == "Day_0" | Treatment == "Non-fumigated chipping grass"  ~ "never",
    Time == "Day_10" & Treatment != "Non-fumigated chipping grass" ~ "recent",
    Time != "Day_0" & Time != "Day_10" & Treatment != "Non-fumigated chipping grass" ~ "past"
  ))
```

## Data Filtering
We filter low abundance OTUs (family level). We can see if those has any affect in our analysis.
```{r}
fam_mask = rowSums(fam.mat) >= 1000
fam.filt = fam.mat[fam_mask,]

fam_mask_harsh = rowSums(fam.mat) >= 10000
fam.filt.harsh = fam.mat[fam_mask_harsh,]
```

## Data Transformation
If we would like to do CLR transformation, we can run the following chunk
However, then we should use Aitchison distance in that case.
Otherwise, we got the following warning message:
`results may be meaningless because data have negative entries in method "bray"`
And also the result might not be correct.
```{r}
#fam.mat = fam.mat + 0.5
#clr_data = clr(fam.mat)
#fam.mat = t(clr_data)
#fam.mat = as.matrix(fam.mat)
#min_val = min(fam.mat[fam.mat>0])
#fam.mat = fam.mat + min_val

# clr_data= apply(fam.mat, 1, function(x) log(x/exp(mean(log(x)))))
# clr_filt = apply(fam.filt, 1, function(x) log(x/exp(mean(log(x)))))
# clr_filt_harsh= apply(fam.filt.harsh, 1, function(x) log(x/exp(mean(log(x)))))
```


## Normalization  
https://microbiomejournal.biomedcentral.com/articles/10.1186/s40168-017-0237-y - This article suggests we normalize (in particular, library sum) before doing PERMANOVA analysis

The following chunk of code is to normalize the data before doing PERMANOVA analysis
In order to use Blue-rays distance, we need to make sure the data is positive, thus using library_sum is the best option.
```{r}
# Log normalize the data would not work because resulting distance matrix would be all 0
# fam.smooth = fam.mat + 0.5
# fam.lr = t( log(t(fam.smooth) / fam.smooth["Unknown_Family",]) )
# fam.lr = fam.lr[rownames(fam.lr)!="Unknown_Family",]

# fam.lr = t(fam.lr)

# Library size normalization
fam.ls = t( t(fam.mat) / colSums(fam.mat) )
fam.ls.filt = t( t(fam.filt) / colSums(fam.filt) )
fam.ls.harsh = t( t(fam.filt.harsh) / colSums(fam.filt.harsh))

fam.ls.nm = fam.ls[-nrow(fam.ls),]
fam.ls.filt.nm = fam.ls.filt[-nrow(fam.ls.filt),]
fam.ls.harsh.nm = fam.ls.harsh[-nrow(fam.ls.harsh),]

fam.mat = t(fam.ls.nm)
fam.mat.filt = t(fam.ls.filt.nm)
fam.mat.harsh = t(fam.ls.harsh.nm)
#fam.mat = as.matrix(fam.mat)
```


## PERMANOVA Analysis

Running Permanova analysis on the 3 dataset to see if fumigation status has any effect on the relative abundance of microbes in the soil.
```{r}
# calculating the distance matrix, we can specify different method depends on our normalization method
dist_matrix_1 = vegdist(fam.mat, method = "bray")
dist_matrix_2 = vegdist(fam.mat.filt, method = "bray")
dist_matrix_3 = vegdist(fam.mat.harsh, method = "bray")

# actually running the permanova analysis
# 1 - all families without filtering
permanova_result_1 = adonis2(dist_matrix_1 ~ fumigation, data = sample_metadata, permutations = 999)
permanova_result_1

# 2 - filtering to only families with abundance >= 1000
permanova_result_2 = adonis2(dist_matrix_2 ~ fumigation, data = sample_metadata, permutations = 999)
permanova_result_2

# 3 - filtering to only families with abundance >= 10000
permanova_result_3 = adonis2(dist_matrix_3 ~ fumigation, data = sample_metadata, permutations = 999)
permanova_result_3
```

## Permanova for Treatment

First, filter out correct data

```{r}
# filter out non-fumigated samples from fam.mat
sample_metadata_treat = sample_metadata %>%
  filter(fumigation != "never")

samp_ids = sample_metadata_treat %>%
  pull(samp_number) %>%
  str_pad(width = 3, pad = "0") %>%
  paste0("Samp", .)

fam.mat.treat = fam.mat[rownames(fam.mat) %in% samp_ids, ]
fam.mat.filt.treat = fam.mat.filt[rownames(fam.mat.filt) %in% samp_ids, ]
fam.mat.harsh.treat = fam.mat.harsh[rownames(fam.mat.harsh) %in% samp_ids, ]
```


Running Permanova analysis to see if treatment has any effect on the relative abundance of microbes in the soil.
```{r}
# calculating the distance matrix, we can specify different method depends on our normalization method
dist_matrix_4 = vegdist(fam.mat.treat, method = "bray")
dist_matrix_5 = vegdist(fam.mat.filt.treat, method = "bray")
dist_matrix_6 = vegdist(fam.mat.harshtreat, method = "bray")

# 1 - all families without filtering
permanova_result_treatment_1 = adonis2(dist_matrix_4 ~ Treatment, data = sample_metadata_treat, permutations = 999)
permanova_result_treatment_1 

# 2 - filtering to only families with abundance >= 1000
permanova_result_treatment_2 = adonis2(dist_matrix_5 ~ Treatment, data = sample_metadata_treat, permutations = 999)
permanova_result_treatment_2 

# 3 - filtering to only families with abundance >= 10000
permanova_result_treatment_3 = adonis2(dist_matrix_6 ~ Treatment, data = sample_metadata_treat, permutations = 999)
permanova_result_treatment_3 
```


## Obervations
By using library sum normalization and bray distance method, the PERMANOVA analysis shows that there is a statistical significant evidence ($P(F^p > F) < 0.001$) in the relative abundance of microbes in the soil is different based on the fumigation status.
However, there is no statistical significant evidence ($P(F^p > F) > 0.8$) in the relative abundance of microbes in the soil is different based on the treatment for all 3 distance matrices.

## Potential Issues and Further Steps
Could try using log normalization and see if the result is different.


## PERMANOVA Analysis for Species OTUs

## Data Preparation
The following chunk of code is to make a species abundance matrix - we only look at the species level.
Note that the "Unknown_Species" value aggregates the counts for all the OTUs whose families are unknown
```{r}
# transpose the data
species.df = bac_abundance %>% 
  mutate(Species = fct_na_value_to_level(Species, "Unknown_Species")) %>% #Missing values included in Unknown_Family level
  group_by(Species) %>%
  summarise(across(starts_with("Samp"), sum))  

#Convert to matrix with species label as rownames 
species = species.df$Species
species.mat = as.matrix(species.df[,-1])
rownames(species.mat) = species

```

No need to try filter due to the large amount of unknown species
## Normalization
Apply library sum normalization to the data
```{r}
## Normalize by library size
species.ls = t( t(species.mat) / colSums(species.mat) )
species.ls.nm = species.ls[-nrow(species.ls),]
species.mat = t(species.ls.nm)
```

## PERMANOVA Analysis
```{r}
## Permanova analysis for fumigation status
dist_matrix_species = vegdist(species.mat, method = "bray")

permanova_result_species = adonis2(dist_matrix_species ~ fumigation, data = sample_metadata, permutations = 999)
permanova_result_species
```

## Permanova for Treatment

```{r}
## Permanova analysis for treatment
species.mat.treat = species.mat[rownames(species.mat) %in% samp_ids, ]
## filter out non-fumigated samples
dist_matrix_species_2 = vegdist(species.mat.treat, method = "bray")

permanova_result_species_2 = adonis2(dist_matrix_species_2 ~ Treatment, data = sample_metadata_treat, permutations = 999)
permanova_result_species_2
```


## Obervations
By using library sum normalization and bray distance method, the PERMANOVA analysis shows that there is a statistical significant evidence  ($P(F^p > F) < 0.001$) in the relative abundance of microbes in the soil is different based on the fumigation status, when looking at species level.
However, there is no statistical significant evidence ($P(F^p > F) = 0.863$) in the relative abundance of microbes in the soil is different based on the treatment, when looking at species level.

## Potential Issues and Further Steps
When looking at species level, there's a huge amount of unknown species, and they were all filtered out during the analysis- so result might not be reliable.

## Overall Observation
The observation is that using current normalization methods and data processing method, we are getting a significant result in PERMANOVA analysis suggesting both the fumigation status has effect on the abundance of OTUs. But treatment has no effect on the abundance of OTUs



```{r include=FALSE}
# ## Log - for personal use
# Here is a log keeping some result from Permanova analysis

# Intepretation of the results:

# Df: degrees of freedom, indicates the levels in the facor minus one. In this case, there are 3 levels in the fumigation factor, so the Df is 3-1 = 2
# SumOfSqs: sum of squares. Vaiance of distance matrix explained by the factor.
# R^2: R-squared, proportion of variance explained by the factor.
# F: F-statistic, value quantifies the ratio of the between-group variance to the within-group variance.
# Pr(>F): p-value, the probability of observing a value of F or greater, assuming that the null hypothesis is true.


# 1. No normalization nor data processing

#             Df SumOfSqs      R2     F Pr(>F)
# fumigation   2   18.835 0.45052 97.16  0.001 ***
# Residual   237   22.972 0.54948
# Total      239   41.806 1.00000

#             Df SumOfSqs      R2      F Pr(>F)
# Treatment   4    4.736 0.11329 7.5063  0.001 ***
# Residual  235   37.070 0.88671
# Total     239   41.806 1.00000

# 2. Library Sum Normalization

#            Df SumOfSqs      R2      F Pr(>F)
# fumigation   2  16.6306 0.64345 213.86  0.001 ***
# Residual   237   9.2152 0.35655
# Total      239  25.8458 1.00000

#              Df SumOfSqs      R2      F Pr(>F)
# Treatment   4   5.0307 0.19464 14.199  0.001 ***
# Residual  235  20.8151 0.80536
# Total     239  25.8458 1.00000
```




Now we perform same analysis on Fungi data

## Permanova for Fungi

## Data Preparation
Read in the cleaned data
```{r}
fung_abundance = readRDS(here("data", "clean", "fung_abundance.RDS"))
samp_metadata = readRDS(here("data", "clean", "samp_metadata.RDS"))
```

```{r}
genus.df = fung_abundance %>% 
  mutate(Genus = fct_na_value_to_level(Genus, "Unknown")) %>%
  group_by(Genus) %>%
  summarise(across(starts_with("Samp"), sum)) 

#Convert to matrix with genus label as rowname
genuses = genus.df$Genus
genus.mat = as.matrix(genus.df[,-1])
rownames(genus.mat) = genuses
```

The following chunk of code is to make a sample metadata dataframe so we have independent variables (sample metadatas such as fumigation status and treatment) in the permanova analysis.
```{r}
sample_metadata = tibble(samp_metadata)
samps = rownames(fam.mat)
fam.mat = t(fam.mat)
rownames(sample_metadata) = samps

sample_metadata = samp_metadata %>%
  mutate(fumigation = case_when(
    Time == "Day_0" | Treatment == "Non-fumigated chipping grass"  ~ "never",
    Time == "Day_10" & Treatment != "Non-fumigated chipping grass" ~ "recent",
    Time != "Day_0" & Time != "Day_10" & Treatment != "Non-fumigated chipping grass" ~ "past"
  ))
```

We create several masks for filtering for different abundances
```{r}
genus_mask = rowSums(genus.mat) >= 100
genus.filt = genus.mat[genus_mask,]

genus_mask_harsh = rowSums(genus.mat) >= 1000
genus.filt.harsh = genus.mat[genus_mask_harsh,]

```

Now we normalize the data
```{r}
genus.ls = t( t(genus.mat) / colSums(genus.mat) )
genus.ls.filt = t( t(genus.filt) / colSums(genus.filt) )
genus.ls.harsh = t( t(genus.filt.harsh) / colSums(genus.filt.harsh))

genus.ls.nm = genus.ls[-nrow(genus.ls),]
genus.ls.filt.nm = genus.ls.filt[-nrow(genus.ls.filt),]
genus.ls.harsh.nm = genus.ls.harsh[-nrow(genus.ls.harsh),]

genus.mat = t(genus.ls.nm)
genus.mat.filt = t(genus.ls.filt.nm)
genus.mat.harsh = t(genus.ls.harsh.nm)
```


## PERMANOVA Analysis

Running Permanova analysis on the 3 dataset to see if fumigation status has any effect on the relative abundance of microbes in the soil.
```{r}
# calculating the distance matrix, we can specify different method depends on our normalization method
dist_matrix_fun_1 = vegdist(genus.mat, method = "bray")
dist_matrix_fun_2 = vegdist(genus.mat.filt, method = "bray")
dist_matrix_fun_3 = vegdist(genus.mat.harsh, method = "bray")

# actually running the permanova analysis
# 1 - all families without filtering
permanova_result_fun_1 = adonis2(dist_matrix_fun_1 ~ fumigation, data = sample_metadata, permutations = 999)
permanova_result_fun_1

# 2 - filtering to only families with abundance >= 1000
permanova_result_fun_2 = adonis2(dist_matrix_fun_2 ~ fumigation, data = sample_metadata, permutations = 999)
permanova_result_fun_2

# 3 - filtering to only families with abundance >= 10000
permanova_result_fun_3 = adonis2(dist_matrix_fun_3 ~ fumigation, data = sample_metadata, permutations = 999)
permanova_result_fun_3
```


## Permanova for Treatment

First, filter out correct data

```{r}
# filter out non-fumigated samples from fam.mat
sample_metadata_treat = sample_metadata %>%
  filter(fumigation != "never")

samp_ids = sample_metadata_treat %>%
  pull(samp_number) %>%
  str_pad(width = 3, pad = "0") %>%
  paste0("Samp", .)

genus.mat.treat = genus.mat[rownames(genus.mat) %in% samp_ids, ]
genus.mat.filt.treat = genus.mat.filt[rownames(genus.mat.filt) %in% samp_ids, ]
genus.mat.harsh.treat = genus.mat.harsh[rownames(genus.mat.harsh) %in% samp_ids, ]
```

Running Permanova analysis to see if treatment has any effect on the relative abundance of microbes in the soil.
```{r}
# calculating the distance matrix, we can specify different method depends on our normalization method
dist_matrix_fun_4 = vegdist(genus.mat.treat, method = "bray")
dist_matrix_fun_5 = vegdist(genus.mat.filt.treat, method = "bray")
dist_matrix_fun_6 = vegdist(genus.mat.harsh.treat, method = "bray")

# 1 - all families without filtering
permanova_result_treatment_fun_1 = adonis2(dist_matrix_fun_4 ~ Treatment, data = sample_metadata_treat, permutations = 999)
permanova_result_treatment_fun_1 

# 2 - filtering to only families with abundance >= 1000
permanova_result_treatment_fun_2 = adonis2(dist_matrix_fun_5 ~ Treatment, data = sample_metadata_treat, permutations = 999)
permanova_result_treatment_fun_2 

# 3 - filtering to only families with abundance >= 10000
permanova_result_treatment_fun_3 = adonis2(dist_matrix_fun_6 ~ Treatment, data = sample_metadata_treat, permutations = 999)
permanova_result_treatment_fun_3 
```


## Permanova for Species

```{r}
# transpose the data
species.df = fung_abundance %>% 
  mutate(Species = fct_na_value_to_level(Species, "Unknown_Species")) %>% #Missing values included in Unknown_Family level
  group_by(Species) %>%
  summarise(across(starts_with("Samp"), sum))  

#Convert to matrix with species label as rownames 
species = species.df$Species
species.mat = as.matrix(species.df[,-1])
rownames(species.mat) = species

```

## Normalization
Apply library sum normalization to the data
```{r}
## Normalize by library size
species.ls = t( t(species.mat) / colSums(species.mat) )
species.ls.nm = species.ls[-nrow(species.ls),]
species.mat = t(species.ls.nm)
```

## PERMANOVA Analysis
```{r}
## Permanova analysis for fumigation status
dist_matrix_species = vegdist(species.mat, method = "bray")

permanova_result_species = adonis2(dist_matrix_species ~ fumigation, data = sample_metadata, permutations = 999)
permanova_result_species
```

## Permanova for Treatment

```{r}
## Permanova analysis for treatment
species.mat.treat = species.mat[rownames(species.mat) %in% samp_ids, ]
## filter out non-fumigated samples
dist_matrix_species_2 = vegdist(species.mat.treat, method = "bray")

permanova_result_species_2 = adonis2(dist_matrix_species_2 ~ Treatment, data = sample_metadata_treat, permutations = 999)
permanova_result_species_2
```

We observe similar pattern in fungi compare to bacteria. Fumigation status has effect on the abundance of OTUs, but treatment does not.