---
title: "Linear Regression on Alpha Diversity and Fumigation"
author: "Tianyi Xu"
date: "2023-05-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(vegan)
library(tidyverse)
library(here)
library(forcats)
library(compositions)
library(dplyr)
library(stringr)
```

## Background

The following code performs linear regression to compare the alpha diversity by fumigation.

## Load Data

```{r}
samp_metadata = readRDS(here("data", "clean", "samp_metadata.RDS"))
bac_abundance = readRDS(here("data", "clean", "bac_abundance.RDS"))
```

The following chunk of code is to make a family abundance dataframe - we only look at the family level first.
Note that the "Unknown_Family" value aggregates the counts for all the OTUs whose families are unknown
```{r}
# transpose the data
fam.df = bac_abundance %>% 
  mutate(Family = fct_na_value_to_level(Family, "Unknown_Family")) %>% #Missing values included in Unknown_Family level
  group_by(Family) %>%
  summarise(across(starts_with("Samp"), sum))  

#Convert to matrix with family label as rownames 
fams = fam.df$Family
fam.mat = as.matrix(fam.df[,-1])
rownames(fam.mat) = fams

fam.mat = t(fam.mat)
```

```{r}

sample_metadata = tibble(samp_metadata)
samps = rownames(fam.mat)
rownames(sample_metadata) = samps

sample_metadata = samp_metadata %>%
  mutate(fumigation = case_when(
    Time == "Day_0" | Treatment == "Non-fumigated chipping grass"  ~ "never",
    Time == "Day_10" & Treatment != "Non-fumigated chipping grass" ~ "recent",
    Time != "Day_0" & Time != "Day_10" & Treatment != "Non-fumigated chipping grass" ~ "past"
  ))
```

## Calculate Alpha Diversity

```{r}
# Calculate alpha diversity (shannon index) for each sample
alpha_div <- diversity(fam.mat, index = "shannon")

sample_metadata = sample_metadata %>%
  mutate(shannon = alpha_div)
```

Here is a plot to visualize the distribution of alpha diversity

```{r}
ggplot(sample_metadata, aes(x=fumigation, y=shannon)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title="Alpha-Diversity by Fumigation Status", x="Fumigation Status", y="Alpha-Diversity")

```

## Linear Regression
First fit the linear regression for alpha diversity and fumigation status.
```{r}
model = lm(shannon ~ fumigation, data = sample_metadata)
summary(model)
```

We check the residual to see if linear model is appropriate
```{r}
residuals = resid(model)
fitted_values = fitted(model)

ggplot() +
  geom_point(aes(x = fitted_values, y = residuals)) +
  geom_hline(yintercept = 0, color = "red") +
  theme_minimal() +
  labs(title="Residual Plot", x="Fitted values", y="Residuals")
```
