---
title: "Linear Regression on Alpha Diversity and Fumigation Using Shannon Index"
author: "Tianyi Xu"
date: "2023-05-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(vegan)
library(tidyverse)
library(here)
library(forcats)
library(compositions)
library(dplyr)
library(stringr)
library(lmtest)
library(sandwich)
library(dunn.test)
```

## Background

The following code performs linear regression to compare the alpha diversity by fumigation using shannon index.

## Load Data

```{r}
samp_metadata = readRDS(here("data", "clean", "samp_metadata.RDS"))
bac_abundance = readRDS(here("data", "clean", "bac_abundance.RDS"))
```

The following chunk of code is to make a family abundance dataframe - we only look at the family level first.
Note that the "Unknown_Family" value aggregates the counts for all the OTUs whose families are unknown
```{r}
# transpose the data
fam.df = bac_abundance %>% 
  mutate(Family = as.character(Family), Genus = as.character(Genus)) %>%
  mutate(Family = ifelse(Family == "Unknown_Family", Genus, Family)) %>%
  mutate(Family = replace_na(Family, "Unknown")) %>% #Missing values included in Unknown_Family level
  group_by(Family) %>%
  summarise(across(starts_with("Samp"), sum))  

#Convert to matrix with family label as rownames 
fams = fam.df$Family
fam.mat = as.matrix(fam.df[,-1])
rownames(fam.mat) = fams
fam.mat = fam.mat[rownames(fam.mat)!="Unknown",]
fam.mat = t(fam.mat)
```

```{r}

sample_metadata = tibble(samp_metadata)
samps = rownames(fam.mat)
rownames(sample_metadata) = samps

sample_metadata = samp_metadata %>%
  mutate(fumigation = case_when(
    Time == "Day_0" | Treatment == "Non-fumigated chipping grass"  ~ "never",
    Time == "Day_10" & Treatment != "Non-fumigated chipping grass" ~ "recent",
    Time != "Day_0" & Time != "Day_10" & Treatment != "Non-fumigated chipping grass" ~ "past"
  ))
```

## Calculate Alpha Diversity

```{r}
# Calculate alpha diversity (shannon index) for each sample
alpha_div = diversity(fam.mat, index = "shannon")

sample_metadata = sample_metadata %>%
  mutate(shannon = alpha_div)
```


## Linear Regression for Fumigation Status and Treatment with Alpha Diversity

```{r}
model_both = lm(shannon ~ fumigation * Treatment, data = sample_metadata)
summary(model_both)
```


Here is 2 models for fumigation and treatment separately. 

## Linear Regression

Here is a plot to visualize the distribution of alpha diversity

```{r}
# anova for fumigation status
anova_result = aov(shannon ~ fumigation, data = sample_metadata)
anova_result
```

```{r}
#TukeyHSD for fumigation status
tukey_result = TukeyHSD(anova_result)
tukey_result
```

Non-parametric test for fumigation status
```{r}
# Kruksal-Wallis test for fumigation status
krusal_result = kruskal.test(shannon ~ fumigation, data = sample_metadata)
krusal_result
```

```{r}
# dunns test for fumigation status
dunn_test_result = dunn.test(sample_metadata$shannon, sample_metadata$fumigation, method = "bonferroni")
dunn_test_result
```

```{r}
# Creating labels for each group
labels_fumigation_b <- c("never" = "a",  # Base group
            "past" = "b",   # Not significantly different from base
            "recent" = "c") # Significantly different from base

```

```{r}
# ggplot(sample_metadata, aes(x=fumigation, y=shannon)) +
#   geom_boxplot() +
#   theme_minimal() +
#   labs(title="Alpha-Diversity (Shannon) by Fumigation Status", x="Fumigation Status", y="Alpha-Diversity (Shannon)")

# Assuming 'sample_metadata' is your data frame
# and 'labels' is your vector of labels created earlier
labels_vector_fumigation_b <- setNames(labels_fumigation_b[unique(sample_metadata$fumigation)], unique(sample_metadata$fumigation))

# Calculate a suitable y position for the labels
max_y <- max(sample_metadata$shannon, na.rm = TRUE) * 1  # 10% above the max to ensure visibility

group_max_fumigation_b = sample_metadata %>%
  group_by(fumigation) %>%
  summarise(max_y = max(shannon, na.rm = TRUE)) %>%
  ungroup()

label_data_fumigation_b = left_join(group_max_fumigation_b, data.frame(fumigation = names(labels_vector_fumigation_b), label = labels_vector_fumigation_b), by = "fumigation")

label_data_fumigation_b$max_y <- label_data_fumigation_b$max_y * 1

# Creating the plot
p_fumigation_b <- ggplot(sample_metadata, aes(x=fumigation, y=shannon)) +
  geom_boxplot() +
  theme_minimal() +
  labs(x = NULL, y="Alpha-Diversity (Shannon)") +
  geom_text(data = label_data_fumigation_b,
            aes(x = fumigation, y = max_y, label = label),
            position = position_dodge(width = 0.75), vjust = -0.5, size = 10) +
            theme(
    plot.title = element_text(size = 25),  # Increase the size of the plot title
    axis.title.x = element_text(size = 30, margin = margin(t = 15)),  # Increase the size of the x-axis title
    axis.title.y = element_text(size = 30, margin = margin(r = 15)),  # Increase the size of the y-axis title
    axis.text.x = element_text(size = 25),  # Increase the size of the x-axis labels
    axis.text.y = element_text(size = 25)   # Increase the size of the y-axis labels
  )

# Print the plot
print(p_fumigation_b)
```

```{r}
# save plot as pdf
ggsave("plot_high_quality/shannon_fumigation_bacteria.pdf", 
       p_fumigation_b,
       width = 14,
       height = 10,
       dpi = 600,
       device = cairo_pdf,
       bg = "white",
       limitsize = FALSE)
```



First fit the linear regression for alpha diversity (shannon) and fumigation status.

```{r}
model = lm(shannon ~ fumigation, data = sample_metadata)
summary(model)
```

Check the residual plot to see if linear model is appropriate

```{r}
residuals = resid(model)
fitted_values = fitted(model)

ggplot() +
  geom_point(aes(x = fitted_values, y = residuals)) +
  geom_hline(yintercept = 0, color = "red") +
  theme_minimal() +
  labs(title="Residual Plot", x="Fitted values", y="Residuals")
```

## Linear Model for Treatment

```{r}
# filter out samples that is not fumigated
sample_metadata_treat = sample_metadata %>%
  filter(fumigation != "never")

```

```{r}
# anova for treatment
anova_result_treat = aov(shannon ~ Treatment, data = sample_metadata_treat)
anova_result_treat
```

```{r}
#TukeyHSD for treatment
tukey_result_treat = TukeyHSD(anova_result_treat)
tukey_result_treat
```

```{r}
# krusal test for treatment
krusal_result_treat = kruskal.test(shannon ~ Treatment, data = sample_metadata_treat)
krusal_result_treat
```

```{r}
# dunns test for treatment
dunn_test_result_treat = dunn.test(sample_metadata_treat$shannon, sample_metadata_treat$Treatment, method = "bonferroni")
dunn_test_result_treat
```

```{r}
# boxplot for alpha diversity by treatment
# ggplot(sample_metadata_treat, aes(x=Treatment, y=shannon)) +
#   geom_boxplot() +
#   theme_minimal() +
#   labs(title="Alpha-Diversity (Shannon) by Treatment", x="Treatment", y="Alpha-Diversity (Shannon)")

# Creating labels for each group
labels_treatment_b <- c("Mirimichi+Biosolids" = "a",  # Base group
            "Mirimichi+Swine" = "a",
            "Non-amended" = "a",
            "NSCC House Brew" = "a")

unique_treatments = unique(sample_metadata_treat$Treatment)
labels_vector_treatment_b <- labels_treatment_b[match(unique_treatments, names(labels_treatment_b))]
names(labels_vector_treatment_b) <- unique_treatments

group_max_treatment_b = sample_metadata_treat %>%
  group_by(Treatment) %>%
  summarise(max_y = max(shannon, na.rm = TRUE)) %>%
  ungroup()

label_data_treatment_b = left_join(group_max_treatment_b, data.frame(Treatment = names(labels_vector_treatment_b), label = labels_vector_treatment_b), by = "Treatment")

label_data_treatment_b$max_y <- label_data_treatment_b$max_y * 1

# Creating the plot
p_treatment_b <- ggplot(sample_metadata_treat, aes(x=Treatment, y=shannon)) +
  geom_boxplot() +
  theme_minimal() +
  labs(x = NULL, y="Alpha-Diversity (Shannon)") +
  geom_text(data = label_data_treatment_b,
            aes(x = Treatment, y = max_y, label = label),
            position = position_dodge(width = 0.75), vjust = -0.5, size = 10) +
            theme(
    plot.title = element_text(size = 25),  # Increase the size of the plot title
    axis.title.x = element_text(size = 30, margin = margin(t = 15)),  # Increase the size of the x-axis title
    axis.title.y = element_text(size = 30, margin = margin(r = 15)),  # Increase the size of the y-axis title
    axis.text.x = element_text(size = 15),  # Increase the size of the x-axis labels
    axis.text.y = element_text(size = 25)   # Increase the size of the y-axis labels
  )


# Print the plot
print(p_treatment_b)
```

```{r}
# save plot as pdf
ggsave("plot_high_quality/shannon_treatment_bacteria.pdf", 
       p_treatment_b,
       width = 14,
       height = 10,
       dpi = 600,
       device = cairo_pdf,
       bg = "white",
       limitsize = FALSE)
```

Linear Model for treatment and alpha diversity (shannon)
```{r}
model_t = lm(shannon ~ Treatment, data = sample_metadata_treat)
summary(model_t)
```

Residual Plot

```{r}
residuals_t = resid(model_t)
fitted_values_t = fitted(model_t)

ggplot() +
  geom_point(aes(x = fitted_values_t, y = residuals_t)) +
  geom_hline(yintercept = 0, color = "red") +
  theme_minimal() +
  labs(title="Residual Plot", x="Fitted values", y="Residuals")
```


It seems linear model model is not appropriate for treatment alone, which is consistent with findings from the permanova analysis.

## Model Assessment using ANOVA

```{r}
# Model assessment for fumigation status

# first, construct a linear model with both fumigation and treatment
model_treat_fum = lm(shannon ~ fumigation + Treatment, data = sample_metadata)

# then, construct a linear model with only fumigation
model_fum = lm(shannon ~ fumigation, data = sample_metadata)

# compare the two models using anova
anova_robust = waldtest(model_fum, model_treat_fum, vcov = vcovHC)
anova_robust
```

```{r}
# Model assessment for treatment

# first, construct a linear model with both fumigation and treatment
model_treat_fum = lm(shannon ~ fumigation + Treatment, data = sample_metadata_treat)

# then, construct a linear model with only treatment
model_treat = lm(shannon ~ Treatment, data = sample_metadata_treat)

# compare the two models using anova
waldtest(model_treat, model_treat_fum, vcov = vcovHC)
```



## Model for Fungi 

Reading in data
```{r}
fung_abundance = readRDS(here("data", "clean", "fung_abundance.RDS"))
samp_metadata = readRDS(here("data", "clean", "samp_metadata.RDS"))
```

Data cleaning
```{r}
genus.df = fung_abundance %>% 
  mutate(Genus = as.character(Genus), Species = as.character(Species)) %>%
  mutate(Genus = ifelse(Genus == "Unknown_Genus", Species, Genus)) %>%
  mutate(Genus = replace_na(Genus, "Unknown")) %>% #Missing values included in Unknown level
  group_by(Genus) %>%
  summarise(across(starts_with("Samp"), sum)) 

genus = genus.df$Genus
genus.mat = as.matrix(genus.df[,-1])
rownames(genus.mat) = genus
genus.mat = genus.mat[rownames(genus.mat)!="Unknown",]
genus.mat = t(genus.mat)

sample_metadata = tibble(samp_metadata)
samps = rownames(genus.mat)
rownames(sample_metadata) = samps

sample_metadata = samp_metadata %>%
  mutate(fumigation = case_when(
    Time == "Day_0" | Treatment == "Non-fumigated chipping grass"  ~ "never",
    Time == "Day_10" & Treatment != "Non-fumigated chipping grass" ~ "recent",
    Time != "Day_0" & Time != "Day_10" & Treatment != "Non-fumigated chipping grass" ~ "past"
  ))
```

Calculating the alpha diversity using shannon index

```{r}
alpha_div_fungi = diversity(genus.mat, index = "shannon")

sample_metadata = sample_metadata %>%
  mutate(shannon_fungi = alpha_div_fungi)
```


## Linear Regression for both the Fumigation Status and Treatment with Alpha Diversity

```{r}
model_both = lm(shannon_fungi ~ fumigation * Treatment, data = sample_metadata)
summary(model_both)
```

Here is 2 models for fumigation and treatment separately. (Since we know treatment itself does not have much impact).

## Linear Regression

Here is a plot to visualize the distribution of alpha diversity using shannon index

```{r}
# anova for fumigation status
anova_result = aov(shannon_fungi ~ fumigation, data = sample_metadata)
anova_result
```

```{r}
#TukeyHSD for fumigation status
tukey_result = TukeyHSD(anova_result)
tukey_result
```

Non-parametric test for fumigation status

```{r}
# Kruksal-Wallis test for fumigation status
krusal_result = kruskal.test(shannon_fungi ~ fumigation, data = sample_metadata)
krusal_result
```

```{r}
# dunns test for fumigation status
dunn_test_result = dunn.test(sample_metadata$shannon_fungi, sample_metadata$fumigation, method = "bonferroni")
dunn_test_result
```

```{r}
# ggplot(sample_metadata, aes(x=fumigation, y=shannon_fungi)) +
#   geom_boxplot() +
#   theme_minimal() +
#   labs(title="Alpha-Diversity (Shannon) for Fungi by Fumigation Status", x="Fumigation Status", y="Alpha-Diversity (Shannon)")

# Creating labels for each group

# Creating labels for each group
labels_fumigation_f <- c("never" = "a",  # Base group
            "past" = "b",   # Not significantly different from base
            "recent" = "b") # Significantly different from base

```

```{r}
labels_vector_fumigation_f <- setNames(labels_fumigation_f[unique(sample_metadata$fumigation)], unique(sample_metadata$fumigation))

# Calculate a suitable y position for the labels
group_max_fumigation_f = sample_metadata %>%
  group_by(fumigation) %>%
  summarise(max_y = max(shannon_fungi, na.rm = TRUE)) %>%
  ungroup()

label_data_fumigation_f = left_join(group_max_fumigation_f, data.frame(fumigation = names(labels_vector_fumigation_f), label = labels_vector_fumigation_f), by = "fumigation")

label_data_fumigation_f$max_y <- label_data_fumigation_f$max_y * 1

# Creating the plot
p_fumigation_f <- ggplot(sample_metadata, aes(x=fumigation, y=shannon_fungi)) +
  geom_boxplot() +
  theme_minimal() +
  labs(x="Fumigation Status", y="Alpha-Diversity (Shannon)") +
  geom_text(data = label_data_fumigation_f,
            aes(x = fumigation, y = max_y, label = label),
            position = position_dodge(width = 0.75), vjust = -0.5, size = 10) + 
           theme(
    plot.title = element_text(size = 25),  # Increase the size of the plot title
    axis.title.x = element_text(size = 30, margin = margin(t = 15)),  # Increase the size of the x-axis title
    axis.title.y = element_text(size = 30, margin = margin(r = 15)),  # Increase the size of the y-axis title
    axis.text.x = element_text(size = 25),  # Increase the size of the x-axis labels
    axis.text.y = element_text(size = 25)   # Increase the size of the y-axis labels
  )
# Print the plot
print(p_fumigation_f)
```

```{r}
# save plot as pdf
ggsave("plot_high_quality/shannon_fumigation_fungi.pdf", 
       p_fumigation_f,
       width = 14,
       height = 10,
       dpi = 600,
       device = cairo_pdf,
       bg = "white",
       limitsize = FALSE)
```



First fit the linear regression for alpha diversity and fumigation status.

```{r}
model = lm(shannon_fungi ~ fumigation, data = sample_metadata)
summary(model)
```

Check the residual to see if linear model is appropriate
```{r}
residuals = resid(model)
fitted_values = fitted(model)

ggplot() +
  geom_point(aes(x = fitted_values, y = residuals)) +
  geom_hline(yintercept = 0, color = "red") +
  theme_minimal() +
  labs(title="Residual Plot", x="Fitted values", y="Residuals")
```

## Linear Model for Treatment

```{r}
# filter out samples that is not fumigated
sample_metadata_treat = sample_metadata %>%
  filter(fumigation != "never")
```

```{r}
# anova for fumigation status
anova_result = aov(shannon_fungi ~ Treatment, data = sample_metadata_treat)
anova_result
```

```{r}
#TukeyHSD for fumigation status
tukey_result = TukeyHSD(anova_result)
tukey_result
```

Non parametric test for treatment

```{r}
# Kruksal-Wallis test for fumigation status
krusal_result = kruskal.test(shannon_fungi ~ Treatment, data = sample_metadata_treat)
krusal_result
```

```{r}
# dunns test for fumigation status
dunn_test_result = dunn.test(sample_metadata_treat$shannon_fungi, sample_metadata_treat$Treatment, method = "bonferroni")
dunn_test_result
```


```{r}
# Creating labels for each group
labels_treatment_f <- c("Mirimichi+Biosolids" = "a",  # Base group
            "Mirimichi+Swine" = "a",
            "Non-amended" = "a",
            "NSCC House Brew" = "a")
```

```{r}
unique_treatments = unique(sample_metadata_treat$Treatment)
labels_vector_treatment_f <- labels_treatment_f[match(unique_treatments, names(labels_treatment_f))]
names(labels_vector_treatment_f) <- unique_treatments

group_max_treatment_f = sample_metadata_treat %>%
  group_by(Treatment) %>%
  summarise(max_y = max(shannon_fungi, na.rm = TRUE)) %>%
  ungroup()

label_data_treatment_f = left_join(group_max_treatment_f, data.frame(Treatment = names(labels_vector_treatment_f), label = labels_vector_treatment_f), by = "Treatment")

label_data_treatment_f$max_y <- label_data_treatment_f$max_y * 1

# Creating the plot
p_treatment_f <- ggplot(sample_metadata_treat, aes(x=Treatment, y=shannon_fungi)) +
  geom_boxplot() +
  theme_minimal() +
  labs(x="Treatment", y="Alpha-Diversity (Shannon)") +
  geom_text(data = label_data_treatment_f,
            aes(x = Treatment, y = max_y, label = label),
            position = position_dodge(width = 0.75), vjust = -0.5, size = 10) +
           theme(
    plot.title = element_text(size = 25),  # Increase the size of the plot title
    axis.title.x = element_text(size = 30, margin = margin(t = 15)),  # Increase the size of the x-axis title
    axis.title.y = element_text(size = 30, margin = margin(r = 15)),  # Increase the size of the y-axis title
    axis.text.x = element_text(size = 15),  # Increase the size of the x-axis labels
    axis.text.y = element_text(size = 25)   # Increase the size of the y-axis labels
  )

# Print the plot
print(p_treatment_f)
```

```{r}
# save plot as pdf
ggsave("plot_high_quality/shannon_treatment_fungi.pdf", 
       p_treatment_f,
       width = 14,
       height = 10,
       dpi = 600,
       device = cairo_pdf,
       bg = "white",
       limitsize = FALSE)
```

# boxplot for alpha diversity by treatment
<!-- ggplot(sample_metadata_treat, aes(x=Treatment, y=shannon_fungi)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title="Alpha-Diversity (Shannon) by Treatment", x="Treatment", y="Alpha-Diversity (Shannon)")
``` -->

Linear Model for treatment and alpha diversity

```{r}
model_t = lm(shannon_fungi ~ Treatment, data = sample_metadata_treat)
summary(model_t)
```

Residual Plot

```{r}
residuals_t = resid(model_t)
fitted_values_t = fitted(model_t)

ggplot() +
  geom_point(aes(x = fitted_values_t, y = residuals_t)) +
  geom_hline(yintercept = 0, color = "red") +
  theme_minimal() +
  labs(title="Residual Plot", x="Fitted values", y="Residuals")
```

It seems linear model model is not appropriate for treatment alone, which is consistent with findings from the permanova analysis.

## Model Assessment using ANOVA

```{r}
# Model assessment for fumigation status

# first, construct a linear model with both fumigation and treatment
model_treat_fum = lm(shannon_fungi ~ fumigation + Treatment, data = sample_metadata)

# then, construct a linear model with only fumigation
model_fum = lm(shannon_fungi ~ fumigation, data = sample_metadata)

# compare the two models using anova
waldtest(model_fum, model_treat_fum, vcov = vcovHC)
```

```{r}
# Model assessment for treatment

# first, construct a linear model with both fumigation and treatment
model_treat_fum = lm(shannon_fungi ~ fumigation + Treatment, data = sample_metadata_treat)

# then, construct a linear model with only treatment
model_treat = lm(shannon_fungi ~ Treatment, data = sample_metadata_treat)

# compare the two models using anova
waldtest(model_treat, model_treat_fum, vcov = vcovHC)
```
